---
title: "20200306 Checking CDR3 mutation data validity"
output: html_document
---

```{r}
library(tidyverse)
```

## Overview

Sachin pointed out the other day that some of the mutation data on the CDR3 assessment (specifically the mutations and hotspot mutations) didn't seem to line up. I wanted to check some of this. Here is the .csv that I sent him to analyze (from the Ghosn lab BM3 data)

```{r}

processed_data <- read_csv("/Users/matt/Desktop/TotalSeq/totalseq_repository/test_data/imgt_vdj_outs/summary_files/lineage_processed_data.csv")

```

Now let's pull the CDR3 data which is what he's interested in

```{r}

cdr3_columns <- which(grepl("cdr3", names(processed_data)))

cdr3_data <- processed_data %>% select(1:20, all_of(cdr3_columns))

```

Ok, now let's look at the distribution of the lengths 

```{r}

ggplot(cdr3_data, aes(x = cdr3_aa_length)) + geom_density()
ggplot(cdr3_data, aes(x = nt_cdr3_length)) + geom_density()

summary(cdr3_data$nt_cdr3_length / cdr3_data$cdr3_aa_length)

```

Ok, so the lengths look good. Everything seems to be a triplicate nt 

Now let's look at the number of mutations in the CDR3s.

```{r}

ggplot(cdr3_data, aes(x = cdr3_imgt_nb_of_mutations)) + geom_density()
ggplot(cdr3_data, aes(x = cdr3_imgt_nb_of_aa_changes)) + geom_density()

```

Those both seem reasonable. Let's make sure there aren't more mutations than positions.

```{r}

ggplot(cdr3_data, aes(x = cdr3_imgt_nb_of_mutations / nt_cdr3_length)) + geom_density()
ggplot(cdr3_data, aes(x = cdr3_imgt_nb_of_aa_changes / cdr3_aa_length)) + geom_density()

```

That all looks fine. Low let's look at hotspots and hotspot mutations.

```{r}

summary(cdr3_data$cdr3_hotspot_loci)
summary(cdr3_data$cdr3_hotspot_muts)


```

On face value that seems fine.

```{r}

summary(cdr3_data$cdr3_hotspot_muts / cdr3_data$cdr3_hotspot_loci)

ggplot(cdr3_data, aes(x = cdr3_data$cdr3_hotspot_muts / cdr3_data$cdr3_hotspot_loci)) + geom_density()

sum(cdr3_data$cdr3_hotspot_muts >= cdr3_data$cdr3_hotspot_loci)

cdr3_heavy_chains <- cdr3_data %>% filter(chain_type == "igh")

ggplot(cdr3_heavy_chains, aes(x = cdr3_heavy_chains$cdr3_hotspot_muts / cdr3_heavy_chains$cdr3_hotspot_loci)) + geom_density()

sum(cdr3_heavy_chains$cdr3_hotspot_muts > cdr3_heavy_chains$cdr3_hotspot_loci)

```

So there are 7 sequences that are showing a greater number of hotspot mutations than the number of hotspots.

Now let's look to see if there are instances where the number of hotspot mutations exceeds the number of mutations

```{r}

summary(cdr3_heavy_chains$cdr3_hotspot_muts / cdr3_heavy_chains$cdr3_imgt_nb_of_mutations)

ggplot(cdr3_heavy_chains, aes(x = cdr3_hotspot_muts / cdr3_imgt_nb_of_mutations)) + geom_density() + geom_rug()

sum(cdr3_heavy_chains$cdr3_hotspot_muts > cdr3_heavy_chains$cdr3_imgt_nb_of_mutations, na.rm = TRUE)

hot_muts_exceed_muts <- cdr3_heavy_chains %>% filter(cdr3_hotspot_muts > cdr3_imgt_nb_of_mutations)

head(hot_muts_exceed_muts$indexed_contig_id)

example1 <- hot_muts_exceed_muts %>% filter(indexed_contig_id == "AAAGATGGTACCGTAT-32_contig_1") 
example2 <- hot_muts_exceed_muts %>% filter(indexed_contig_id == "AAAGTAGAGACAGAGA-32_contig_1") 
example3 <- hot_muts_exceed_muts %>% filter(indexed_contig_id == "AACCATGAGGCCCGTT-32_contig_2") 
example4 <- hot_muts_exceed_muts %>% filter(indexed_contig_id == "AACTCTTTCACCACCT-32_contig_1") 
example5 <- hot_muts_exceed_muts %>% filter(indexed_contig_id == "AAGCCGCGTCGGCATC-32_contig_1") 
example6 <- hot_muts_exceed_muts %>% filter(indexed_contig_id == "AAGGCAGCACGGTGTC-32_contig_2") 

paste(example1$nt_cdr3_length, example1$cdr3_imgt_nb_of_mutations, example1$cdr3_hotspot_loci, example1$cdr3_hotspot_muts, sep = ";")
paste(example2$nt_cdr3_length, example2$cdr3_imgt_nb_of_mutations, example2$cdr3_hotspot_loci, example2$cdr3_hotspot_muts, sep = ";")
paste(example3$nt_cdr3_length, example3$cdr3_imgt_nb_of_mutations, example3$cdr3_hotspot_loci, example3$cdr3_hotspot_muts, sep = ";")
paste(example4$nt_cdr3_length, example4$cdr3_imgt_nb_of_mutations, example4$cdr3_hotspot_loci, example4$cdr3_hotspot_muts, sep = ";")
paste(example5$nt_cdr3_length, example5$cdr3_imgt_nb_of_mutations, example5$cdr3_hotspot_loci, example5$cdr3_hotspot_muts, sep = ";")
paste(example6$nt_cdr3_length, example6$cdr3_imgt_nb_of_mutations, example6$cdr3_hotspot_loci, example6$cdr3_hotspot_muts, sep = ";")

```

```{r}

ggplot(hot_muts_exceed_muts, aes(x = nt_cdr3_length)) + geom_density()

```

Well I just looked into the IMGT analysis, and it looks like there is an internal issue where the v_mutation analysis is incorrectly referencing the aligned CDR3 sequence. I have reached out to them for clarification.




























































